---
title: "Netflix Shows And Movies Analysis"
author: "Caleb Mogyabiyedom"
editor: visual
---

## Introduction

Netflix began as a DVD rental service and evolved into the world's largest streaming platform. This analysis explores how Netflix's content library has changed between 2008 and 2021, specifically how the balance between movies and TV shows has shifted, how rapidly the platform scaled its catalogue, and which genres drove that growth. Understanding these trends reveals the strategic decisions Netflix made to dominate the streaming wars and retain subscribers in an increasingly competitive market.

**Key question: How has Netflix's content strategy evolved over time, and what does that tell us about how the platform competes?**

```{r}
#| warning: false
#| message: false
#| echo: false

suppressPackageStartupMessages(library(tidyverse))
library(ggplot2)
netflix_titles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-20/netflix_titles.csv',
show_col_types = FALSE)



netflix_titles_updated <- netflix_titles |>
  mutate(year_added = str_extract(date_added, "\\d{4}"),
         year_added = str_replace_na(year_added, "Unknown"))|>
  mutate(type = str_to_upper(type))|>
  
  filter(year_added != "Unknown")


#view(netflix_titles_updated)


proportion_data <- netflix_titles_updated |>
  group_by(year_added, type) |>
  summarise(count = n(), .groups = "drop") |>
  ungroup() |>
  group_by(year_added) |>
  mutate(proportion = count / sum(count))

#proportion_data



proportion_plot_1 <- ggplot(proportion_data, aes(x = year_added, y = proportion, fill = type)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = c("MOVIE" = "#E50914", "TV SHOW" = "#221F1F")) +
  labs(title = "Proportion of TV Shows and Movies by Year",
       x = "Year Added",
       y = "Proportion",
       fill = "Type") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1))

proportion_plot_1


```

The first plot, "Proportion of TV Shows and Movies by Year", shows the proportion of TV shows and movies added to Netflix each year. This is a stacked bar chart, where each year is represented by a bar, with segments showing the proportion of TV shows and movies.In the graph, movies tended to dominate the viewing options especially from 2009 to 2012. Although TV Shows in 2008 had the same as movies. In recent years however,the proportion of TV shows have increased with about 25 percent of viewing options being TV shows.

```{r}
#| echo: false
trend_plot <- netflix_titles_updated |>
  count(year_added, type) |>
  ggplot(aes(x = year_added, y = n, color = type, group = type)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2.5) +
  scale_color_manual(values = c("MOVIE" = "#E50914", "TV SHOW" = "#221F1F")) +
  labs(title = "Trend of TV Shows and Movies Added by Year",
       x = "Year Added",
       y = "Count",
       color = "Type") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1))

trend_plot



```

The second plot, "Trend of TV Shows and Movies Added by Year", shows the actual counts of TV shows and movies added per year. This is a line plot with different colors representing movies and TV shows. Between 2015 and 2020, they both experienced steep growths however movies added far outweighed TV Shows added

```{r}
#| echo: false
genres <- c("Thriller", "Action", "Dramas", "Romantic", "Horror", "Crime", "Sci-Fi", "Fantasy", "Comedies", "Documentaries", "Kids\' TV")

genre<- netflix_titles_updated|>
  mutate(is_genre = str_detect(listed_in, str_c(genres, collapse = "|")))|>
  mutate(genre_label = str_extract(listed_in, str_c(genres, collapse = "|")))|>
  group_by(year_added, type, genre_label) |>
  summarise(count = n(), .groups = "drop")
#genre

 
ggplot(genre, aes(x = year_added, y = count, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("MOVIE" = "#E50914", "TV SHOW" = "#221F1F")) +
  facet_wrap(~ genre_label) +
  labs(title = "Variations in Number of Movies and TV Shows by Genre and Year",
       x = "Year Added",
       y = "Count of Titles",
       fill = "Type") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 14),
        axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(face = "bold"))
```

The third plot, "Variations in Number of Movies and TV Shows by Genre and Year", shows how different genres have varied over the years with respect to both TV Shows and Movies. In recent years, there's been growth in all the genres, however Comedies, Dramas, Documentary and Action movies rose more highly. Although Crimes, Kid's TV and Romantic movies have all risen in the TV Shows category, they are still below that of movies.

## Conclusion

Taken together, these three analyses tell a clear story: Netflix deliberately transformed itself from a movie-first platform into a broad content powerhouse. In its early years, movies made up the vast majority of the catalogue, but from 2015 onward, Netflix aggressively expanded both volume and variety. The surge in TV shows is particularly telling; series keep subscribers engaged month after month in a way that films cannot. Meanwhile, the genre breakdown reveals that this wasn't random growth, Dramas, Comedies, and Documentaries led the expansion, reflecting Netflix's bet on prestige content that attracts awards attention and cultural conversation. The data suggests Netflix's content strategy has always been intentional, data-driven, and evolving.

**Reference:**

Harmon, J. (n.d.). Tidytuesday/data/2021/2021-04-20/readme.md at main · rfordatascience/tidytuesday. GitHub. https://github.com/rfordatascience/tidytuesday/blob/main/data/2021/2021-04-20/readme.md

The data comes from Kaggle with credit to Shivam Bansal and it was collected from Flixable which is a third-party Netflix search engine.

<https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-04-20/readme.md>